Fast and Scalable Mixed Precision Euclidean Distance Calculations Using GPU Tensor Cores

### 2025-06-19 Paper/main.tex
Fixed GHz capitalizations.

While analyzing our results from Table~\ref{tab:profiler_results}, we discovered a discrepancy between the TC Pipe Utilization metric and our derived TFLOPS. The profiler reported 64\% utilization while, by our calculations, we only achieved 49\%. The explanation for this discrepancy is that at $d{=}4096$, the 250~W power budget for our PCIe A100 is exceeded, and the clock is throttled to 1.12 Ghz. The implications of this finding are discussed in the conclusion.

\begin{table}[!t]
\caption{Nvidia Nsight Compute profiler results on the \synth dataset with $|D|{=}10^5$. We examine $d\in\{128, 256, 4096\}$ to illustrate how key performance metrics of the brute force TC algorithms (\ouralg and \tedjoinbrute) vary as a function of dimensionality. OOM refers to out of shared memory.}
\footnotesize
    \centering
        \begin{tabularx}{\columnwidth}{L|r|r|r|r|r|r} 
        \hline
           \rowcolor{blue!40!white} Metric &\multicolumn{3}{c|}{\ouralg}& \multicolumn{3}{c}{\tedjoinbrute}\\ \hline%\cline{2-6}
            \rowcolor{blue!40!white}  Dimensionality ($d$)& \cellcolor{blue!40!white}$128$ & \cellcolor{blue!40!white}$256$ &\cellcolor{blue!40!white}$4096$ & \cellcolor{blue!40!white}$128$ & \cellcolor{blue!40!white}$256$&\cellcolor{blue!40!white}$4096$ \\ \hline
             DRAM Throughput (\%) & 1.98 & 3.54 & 16.0 &   0.04 & 0.04 & OOM\\ \hline
             SMEM Throughput (\%) &6.49 & 10.5 & 36.1 &   42.3 & 16.0 & OOM\\ \hline
             Bank Conflicts (\%) & 0.00 & 0.00 & 0.00 & 92.3 & 75.0&OOM \\ \hline
             L2 Hit Rate (\%) & 89.8 & 89.6 &  84.4 &  98.9 & 98.9& OOM\\ \hline
             TC Pipe Utilization FP16-32 (\%) &  10.1& 17.8 & 64.0&   N/A & N/A&OOM \\ \hline
             TC Pipe Utilization FP64 (\%) &  N/A & N/A & N/A & 5.75& 1.99 &OOM\\ \hline
             Clock Speed (Ghz) &  1.37 & 1.40 & 1.12 & 1.40 & 1.41 &OOM\\ \hline
        \end{tabularx}
    \label{tab:profiler_results}
\end{table}

>>>>

While analyzing our results from Table~\ref{tab:profiler_results}, we discovered a discrepancy between the TC Pipe Utilization metric and our derived TFLOPS. The profiler reported 64\% utilization while, by our calculations, we only achieved 49\%. The explanation for this discrepancy is that at $d{=}4096$, the 250~W power budget for our PCIe A100 is exceeded, and the clock is throttled to 1.12 GHz. The implications of this finding are discussed in the conclusion.

\begin{table}[!t]
\caption{Nvidia Nsight Compute profiler results on the \synth dataset with $|D|{=}10^5$. We examine $d\in\{128, 256, 4096\}$ to illustrate how key performance metrics of the brute force TC algorithms (\ouralg and \tedjoinbrute) vary as a function of dimensionality. OOM refers to out of shared memory.}
\footnotesize
    \centering
        \begin{tabularx}{\columnwidth}{L|r|r|r|r|r|r} 
        \hline
           \rowcolor{blue!40!white} Metric &\multicolumn{3}{c|}{\ouralg}& \multicolumn{3}{c}{\tedjoinbrute}\\ \hline%\cline{2-6}
            \rowcolor{blue!40!white}  Dimensionality ($d$)& \cellcolor{blue!40!white}$128$ & \cellcolor{blue!40!white}$256$ &\cellcolor{blue!40!white}$4096$ & \cellcolor{blue!40!white}$128$ & \cellcolor{blue!40!white}$256$&\cellcolor{blue!40!white}$4096$ \\ \hline
             DRAM Throughput (\%) & 1.98 & 3.54 & 16.0 &   0.04 & 0.04 & OOM\\ \hline
             SMEM Throughput (\%) &6.49 & 10.5 & 36.1 &   42.3 & 16.0 & OOM\\ \hline
             Bank Conflicts (\%) & 0.00 & 0.00 & 0.00 & 92.3 & 75.0&OOM \\ \hline
             L2 Hit Rate (\%) & 89.8 & 89.6 &  84.4 &  98.9 & 98.9& OOM\\ \hline
             TC Pipe Utilization FP16-32 (\%) &  10.1& 17.8 & 64.0&   N/A & N/A&OOM \\ \hline
             TC Pipe Utilization FP64 (\%) &  N/A & N/A & N/A & 5.75& 1.99 &OOM\\ \hline
             Clock Speed (GHz) &  1.37 & 1.40 & 1.12 & 1.40 & 1.41 &OOM\\ \hline
        \end{tabularx}
    \label{tab:profiler_results}
\end{table}
